Value-Suppressing Uncertainty Maps

### 2017-09-17 paper/main.tex
Created exp1 figure
Fresh commit to deal with paper weirdness.
Jeff's revisions.

In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}. However, even simple bivariate maps can be difficult for a general audience to interpret: due to their additional complexity, Wainer \& Francolini~\cite{wainer1980empirical} reported high levels of error for participants even for ``elementary'' graph reading tasks, and additionally that bivariate legends were more difficult to memorize and internalize than their univariate equivalents. Therefore, bivariate maps in practice are often limited to a small set of output colors (say, a 4x4 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}. For VSUPs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly.

Given that bivariate maps, for reasons of either practicality or perceptual fidelity, have only a limited budget of outputs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly. Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty, where, frequently, uncertain information ought to have less valence on the decision-making process, or at least less prominence in the analysis, than certain information.

@@

However, if both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values. VSUPs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUP will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low. VSUPs therefore act as both a filtering mechanism (in that values with too much uncertainty are mapped to the same glyph), as well as an implicit test of effect size (in that smaller and smaller changes in value are visible as uncertainty decreases). This strategy of dampening low quality or highly uncertain values in maps in order to focus on more informative regions has measurable benefits, including the removal of statistically spurious visual patterns, and the highlighting of regions of interest~\cite{correll2017surprise}.

@@

There are multiple choices that designers must make before creating a VSUP. Specifically, they must choose (1) which visual channels map to value (1), uncertainty (2), and the structure of the underlying VSUP tree (3).

@@

When selecting a visual channel to represent value, the designer should consider the perceptual integrality of the value and uncertainty visual channels. For instance, we use the Viridis~\cite{viridis} color map frequently in this paper, as many other standard color ramps (such as sequential ColorBrewer~\cite{harrower2003colorbrewer} ramps), following usual best practices, interpolate in both hue and value~\cite{ware1988color}. This interpolation in value interferes with our uncertainty encoding, introducing ambiguity. Ramps with similar saturation and value among colors, such as Virids, are less ambiguous.

@@

We present a set of examples showing how VSUPs can be applied to realworld datasets. The decision of VSUPs to allocate color bins asymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.

@@

Both maps illustrate a similar temporal trend: flight delays are shorter towards the beginning of the day, increasing on average over time. The traditional bivariate map affords only a coarse examination of this trend: two visible ``blocks'' of color, early in the day, and later in the afternoon. The VSUP, in contrast, has sufficient color resolution to show a quasi-linear trend. Uncertainty is high only for a few bins for which there are small number of flights, either on the weekends or late at night. The VSUP does not obscure any significant trends in the high uncertainty data, despite having fewer color categories overall.

@@

Using a dataset from a group of virologists interested in the population dynamics of viral mutation~\cite{o2012conditional}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV). Next generation sequencing (NGS) allows biologists to collect and analyze large amounts of genomics data. However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs. In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but to not be distracted by false positives. The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots. Extending the ``confidence fog'' metaphor used in viral genome browsers like LayerCake~\cite{correll2015layercake}, VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while not being distracted by noise. Smaller scale ``hot spots'' that are mapped to the same magenta color in the traditional map are visible as bright orange lines. Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) that could be plausibly considered potential regions of interest in the 2D map, are encoded as highly uncertain and of ambiguous value in the VSUPs. In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the  stronger signals earlier in the genome.

\textbf{A few thoughts about this section based on the current figures. (1) Fig. 3 in particular has two graphics that look pretty similar. Is there anything we should do to draw attention to the most salient differences between conditions? At minimum, we have the text, but we could consider annotations or even a third vis that is a heatmap of differences between charts. (2) A third example using a different visual form might be useful. For example, a node-link network with value + uncertainty associated with each node.}

@@

We performed two crowdsourced experiments on Amazon's Mechanical Turk to evaluate the effectiveness of VSUPs for integrating uncertainty and value information in visualizations. This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision. The experiments consisted of two separate tasks:

\begin{enumerate}
	\item An experiment with an \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions. E.g., ``click on the region of the chart with a value of $0.1$ and an uncertainty of $0.2$.''
	\item An experiment with a \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value. E.g., ``place your $5$ ships on safe locations on the board.''
\end{enumerate}

We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%. As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test. Based on piloting, we paid participants \$2 dollars for participation, for a target rate of \$8/hour. After completion of the main tasks, we solicited demographic information, including a risk aversion assay from Mandrik \& Bao~\cite{mandrik2005exploring}. We recruited a total of $48$ participants for this experiment ($24$ for each experiment): $24$ female, $24$ male, ($M_{\text{age}}$= $34$, $SD_{\text{age}}$ = $9.1$). Participants from the first experiment were excluded from participating in the second.

@@

For the identification task, we gave participants a $5$x$5$ heatmap, and asked them to click on a region of the heatmap that encodes a particular value,uncertainty pair. To support comparability across conditions, we chose a set of target pairs such that each target mapped to a unique color across all encoding types. This resulted in (4+4+2+1)=11 valid targets. For each trial, these 11 targets were placed randomly in a heatmap; the remaining 13 cells of the heatmap were distractors randomly sampled from this target list.

@@

We had two initial hypotheses, based on prior work on bivariate maps:
\begin{enumerate}
	\item Juxtaposed maps, by introducing a second search task to the identification task (searching for the proper values in two, rather than one map), would have poorer performance than the other conditions.
	\item Continuous maps required more specific and accurate color matching between value and color, and so would have poorer performance than the other conditions.
\end{enumerate}

@@

\subsubsection{Results}

We performed a repeated measures ANOVA on our results to measure the effect of juxtaposed versus superimposed maps, and continuous versus discrete legends, on accuracy. Participant ID was included as a random effect. Figure %TODO
shows the performance of each of our 8 conditions.

Our results support our first hypothesis. Juxtaposition was a significant effect ($F(1,166)=5.5$,$p=0.02$). A post-hoc t-test confirmed ($t=2.5$, $p=0.01$)that superimposed charts ($M=0.58$, $SD=0.5$) performed significantly better than juxtaposed ($M=0.51$, $SD=0.49$) ones. For our identification task, juxtaposition requires at least two search tasks (one to locate squares with the correct data value, another to locate ones with the correct uncertainty value), which introduces additional error compared to conditions where value and uncertainty are displayed simultaneously.

Our results also support our second hypothesis. Discretization was a significant effect ($F(1,166)=30$,$p<0.01$). A post-hoc t-test confirmed ($t=6.1$, $p<0.01$) that charts with discrete bins ($M=0.63$, $SD=0.48$) performed significantly better than charts with continuous color maps ($M=0.47$, $SD=0.5$). The lack of quantization bias in continuous maps is countered by the perceptual error in precisely estimating value from color. Relying on a discrete set of output colors simplifies this task.

We performed a second ANOVA amongst just the superimposed, discrete charts, to determine the effect of legend shape (wedge or square) and quantization scheme (VSUP or standard) on performance, with participant ID as a random factor. Neither the legend shape ($F(1,70)=0.03$, $p=0.084$) nor the quantization scheme ($F(1,70)=1.4$, $p=0.24$) were significant effects.

@@

For the prediction task, we gave participants the rules of a game like Battleship. Greis et al. \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience. In our task, the participant and a (fictional) adversary have to place tokens representing ships on a $5$x$5$ spatial grid, with the expectation that certain squares will be hit by missiles. Players have to place all their tokens before continuing. The objective is to minimize the number of your own ships that are hit. In our task, participants were given a map representing the predictions of missile strikes in each location on the grid. The \emph{value} component of the prediction was the likelihood of a missile strike in a particular region: the ship's \emph{safety} if placed on the square. The \emph{uncertainty} component was the confidence in this prediction. Other studies of uncertainty representation, such as in Cox et al.~\cite{cox2013visualizing}, have used "prediction + prediction uncertainty" stimuli to elicit differences in decision-making between visualizations of uncertainty.

@@

Our results from the prior study indicate indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and arc bivariate maps. These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or arc legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli for each of our 24 participants. \figref{fig:conditions} shows these factor levels. Prior to the main task, we included a short replication of the identification task from the prior experiment (with 12 stimuli) for training purposes. Poor performance on this training was used for exclusion purposes. Three people with low accuracy ($6\%$, $6\%$, and $25\%$ accuracy compared to a mean of $70\%$ ) were excluded.

@@

We had two hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and also encourage \emph{caution} by highlighting the ambiguity or untrustworthiness introduced by uncertain data. In particular:
\begin{enumerate}
	\item Participants would choose targets with \textbf{lower uncertainty} in the prediction task when using a VSUP. 
	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value} when using a VSUP.
\end{enumerate}

@@

\subsubsection{Results} 
\taskTwoViolin
Consistent with our first experiment, we found no significant effect of legend shape on either uncertainty ($F(1,61) = 0.01$, $p=0.92$) or value ($(1,61)=3.1$, $p=0.08$) of guesses. This seems to indicate that both wedge- or square-shaped legends promote similar patterns of decision-making. In this paper, and in general, we prefer to employ the wedge-shaped legend for VSUPs, and the square-shaped legend for traditional maps, as it makes the conceptual differences between the two more apparent.

Our results failed to support our first hypothesis. We performed a repeated measures ANOVA on our results to measure the effect of VSUP versus standard quantizations, and square- versus wedge-shaped legends, on average uncertainty in bets. We found no significant effect of quantization scheme on average guess uncertainty ($F(1,61)=0.05$, $p=0.83$). This indicates that there does not appear to be a uniform pattern of risk aversion between the two scale types, as we had surmised.

Our results supported our second hypothesis. A similar ANOVA on average danger value of guesses found that quantization scheme was a significant effect ($F(1,61)=17$, $p<0.01$). A post-hoc t-test confirmed ($t=2.3$, $p=0.02$) that guesses made with VSUPs had significantly higher average danger ($M=0.32$, $SD=0.17$) (and so lower expected value) than traditional maps ($M=0.29$, $SD=0.18$). Since there was no significant difference in average guess uncertainty, and the distribution of danger and uncertainty was identical across conditions, this seems to indicate that participants were using VSUPs moderating their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty. \figref{fig:taskTwoViolin} shows the comparative distribution of uncertainty in participant guesses.

@@

The results from our initial evaluation of VSUPs are promising, showing that even a general audience can make use of uncertainty information in a reasonable way. The way that uncertainty information is presented can have a measurable impact on decision-making. The VSUP strategy of only ``showing'' differences when uncertainty is sufficiently low is somewhat analogous to the inferential statistics such as effects tests. VSUPs might therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.

@@

Often, uncertainty, data quality, or confidence are considered separate information from the data itself, relegated to tooltips or visually distant supplemental charts. We believe, in contrast, that uncertainty information ought to be directly integrated with charts of data itself. This integration introduces additional complexity in the design and presentation of data. Value-Suppressing Uncertainty Palettes represent one strategy for dealing with this complexity, by assigning marks properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high. This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.

>>>>

In general, the quality of bivariate color maps is a multivariate measure involving consideration of not just the component color channels, but also the interpolation scheme and the color of the surround~\cite{bernard2015survey}. However, even simple bivariate maps can be difficult for a general audience to interpret: due to their additional complexity, Wainer \& Francolini~\cite{wainer1980empirical} reported high levels of error for participants even for ``elementary'' graph reading tasks, and additionally that bivariate legends were more difficult to memorize and internalize than their univariate equivalents. Therefore, bivariate maps in practice are often limited to a small set of output colors (say, a 4x4 matrix as in \figref{fig:example}) \cite{robertson1986generation,trumbo1981theory}.

Given that bivariate maps, for reasons of either practicality or perceptual fidelity, have only a limited budget of outputs, we adapt an insight from Dunn~\cite{dunn1989dynamic} that these limited categories ought to be assigned with regards to their importance to analysis tasks, rather than uniformly. Our Value-Suppressing Uncertainty Palettes apply this insight to data with uncertainty, where, frequently, uncertain information ought to be weighted less heavily in a decision-making process, or at least less prominence in an analysis, than certain information.

@@

If both of these assumptions hold, then it follows that the designer of a bivariate map should allocate more mark types to certain values, and fewer mark types to uncertain values. VSUPs codify this decision by \emph{reducing the number of mark categories for representing value as uncertainty increases.} This means that data encoded using a VSUP will make visible only the largest of differences in uncertain data, but highlight comparatively small differences in value when uncertainty is low. VSUPs therefore act as both a filtering mechanism (in that values with too much uncertainty are mapped to the same glyph), as well as an implicit test of effect size (in that smaller and smaller changes in value are visible as uncertainty decreases). This strategy of dampening low quality or highly uncertain values in maps in order to focus on more informative regions has measurable benefits, including the removal of statistically spurious visual patterns, and the highlighting of regions of interest~\cite{correll2017surprise}.

@@

There are multiple choices that designers must make before creating a VSUP. Specifically, they must choose which visual channels map to value (1), uncertainty (2), and the structure of the underlying VSUP tree (3).

@@

When selecting a visual channel to represent value, the designer should consider the perceptual integrality of the value and uncertainty visual channels. For instance, we use the Viridis~\cite{viridis} color map frequently in this paper, as many other standard color ramps (such as sequential ColorBrewer~\cite{harrower2003colorbrewer} ramps), following usual best practices, interpolate in both hue and value~\cite{ware1988color}. This interpolation in value interferes with our uncertainty encoding, introducing ambiguity. Ramps with similar saturation and value among colors, such as Viridis, are less ambiguous.

@@

We present a set of examples showing how VSUPs can be applied to real-world datasets. The decision of VSUPs to allocate color bins asymmetrically amongst value and uncertainty regions affords greater discriminability in regions of the chart that warrant closer scrutiny, while discouraging exploration of regions with noisy or unreliable signals.

@@

Both maps illustrate a similar trend: flight delays are shorter towards the beginning of the day, increasing on average over time. The traditional bivariate map affords only a coarse examination of this trend: two visible ``blocks'' of color, early in the day, and later in the afternoon. The VSUP, in contrast, has sufficient color resolution to show a quasi-linear trend. Uncertainty is high only for a few bins for which there are small number of flights, either on the weekends or late at night. The VSUP does not obscure any significant trends in the high uncertainty data, despite having fewer color categories overall.

@@

Next generation sequencing allows biologists to collect and analyze large amounts of genomics data. However, these techniques often create data quality issues, as they must align large numbers of potentially ambiguous ``reads'' of relatively small numbers of base pairs. Using a dataset from a group of virologists interested in the population dynamics of viral mutation~\cite{o2012conditional}, we visualized the variability in viral genomes across 25 populations of the simian immunodeficiency virus (SIV). In the dataset here, it is important not only to identify hotspots in the viral genome (locations with high rates of mutation or variability), but also to not be distracted by false positives. The traditional bivariate map (\figref{fig:viral2d}) affords the analysis of only a few, large hot spots.

VSUPs (\figref{fig:viralVsum}) enable biologists to explore interesting patterns while deemphasizing noise. Smaller scale ``hot spots'' that are mapped to the same magenta color in the traditional map are visible as bright orange lines. Pink hot spots towards the end of the genome (around reference nucleotide $9,500$) could be considered potential regions of interest in the 2D map; in the VSUP these are encoded as highly uncertain and of ambiguous value. In fact, these regions have systematically low coverage; the lack of reliable data for these regions should discourage analysts from giving their apparent pattern equal consideration from the stronger signals earlier in the genome.

\textbf{A third example using a different visual form might be useful. For example, a node-link network with value + uncertainty associated with each node.}

@@

We performed two crowdsourced experiments on Amazon's Mechanical Turk to evaluate the effectiveness of VSUPs for integrating uncertainty and value information in visualizations. This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision. Each experiment focused on one of two separate tasks:

\begin{enumerate}
	\item An \textbf{identification} task, where we gave participants charts with value and uncertainty information, and asked them to locate specific regions. E.g., ``click on the region of the chart with a value of $0.1$ and an uncertainty of $0.2$.''
	\item A \textbf{prediction} task, inspired by the game Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value. E.g., ``place your $5$ ships on safe locations on the board.''
\end{enumerate}

We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%. As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates~\cite{hardy1945tests} as a pre-test, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test. Based on piloting, we paid participants \$2 dollars for participation, for a target rate of \$8/hour. After completion of the main tasks, we solicited demographic information, including a risk aversion assay from Mandrik \& Bao~\cite{mandrik2005exploring}. We recruited a total of $48$ participants ($24$ for each experiment): $24$ female, $24$ male, ($M_{\text{age}}$= $34$, $SD_{\text{age}}$ = $9.1$). Participants from the first experiment were excluded from participating in the second.

@@

For the identification task, we gave participants a $5$x$5$ heatmap, and asked them to click on a region of the heatmap that encodes a particular (value, uncertainty) pair. To support comparability across conditions, we chose a set of target pairs such that each target mapped to a unique color across all encoding types. This resulted in (4+4+2+1)=11 valid targets. For each trial, these 11 targets were placed randomly in a heatmap; the remaining 13 cells of the heatmap were distractors randomly sampled from this target list.

@@

We had two initial hypotheses, informed by prior work:

\begin{enumerate}
	\item Juxtaposed maps, by introducing a second search task to the identification task (searching for the proper values in two, rather than one map), would have poorer performance than other conditions.
	\item Continuous maps require more specific and accurate color matching between value and color, and so would have poorer performance than discrete conditions.
\end{enumerate}

@@

\subsubsection{Results}
\taskOneFig

We performed a repeated measures ANOVA on our results to measure the effect of juxtaposed versus superimposed maps, and continuous versus discrete legends, on accuracy. Participant ID was included as a random effect. \figref{fig:taskOne} shows the performance of each of our 8 conditions.

The results support our first hypothesis. Juxtaposition was a significant effect ($F(1,166)=5.5$,$p=0.02$). A post-hoc t-test confirmed ($t=2.5$, $p=0.01$) that superimposed charts ($M=0.58$, $SD=0.5$) performed significantly better than juxtaposed ($M=0.51$, $SD=0.49$) ones. For our identification task, juxtaposition requires at least two search tasks (one to locate squares with the correct data value, another to locate ones with the correct uncertainty value), which introduces additional error compared to conditions where value and uncertainty are displayed simultaneously.

The results also support our second hypothesis. Discretization was a significant effect ($F(1,166)=30$,$p<0.01$). A post-hoc t-test confirmed ($t=6.1$, $p<0.01$) that charts with discrete bins ($M=0.63$, $SD=0.48$) performed significantly better than charts with continuous color maps ($M=0.47$, $SD=0.5$). The lack of quantization bias in continuous maps is countered by the perceptual error in precisely estimating value from color. Relying on a discrete set of output colors simplifies this task.

We performed a second ANOVA among the superimposed discrete charts to determine the effect of legend shape (wedge or square) and quantization scheme (VSUP or standard) on performance, with participant ID as a random factor. We did not find a significant effect for either the legend shape ($F(1,70)=0.03$, $p=0.084$) or the quantization scheme ($F(1,70)=1.4$, $p=0.24$).

@@

For the prediction task, we gave participants the rules of a game similar to Battleship. Greis \ea~\cite{greis2016decision} employ these game-like experimental tasks to assess how different visual designs communicate uncertainty information, which can be abstract or complex, to the general audience. In our task, the participant and a (fictional) adversary have to place tokens representing ships on a $5$x$5$ spatial grid, with the expectation that certain squares will be hit by missiles. Players have to place all their tokens before continuing. The objective is to minimize the number of your own ships that are hit. In our task, participants were given a map representing the predictions of missile strikes in each location on the grid. The \emph{value} component of the prediction was the likelihood of a missile strike in a particular region: the ship's \emph{safety} if placed on the square. The \emph{uncertainty} component was the confidence in this prediction. Other studies of uncertainty representation, such as in Cox et al.~\cite{cox2013visualizing}, have used "prediction + prediction uncertainty" stimuli to elicit differences in decision-making between visualizations of uncertainty.

@@

Our results from the prior study indicate that discrete, non-juxtaposed maps outperformed the other bivariate maps we selected, so we limited our study to only 4 types: square and arc bivariate maps. These in turn were either VSUPs or traditional bivariate maps, for a 2 (square or arc legend) x 2 (VSUP or standard) factorial design, with 6 replications, for a total of 24 stimuli for each of our 24 participants. \figref{fig:conditions} shows these factor levels. Prior to the main task, we included a short replication of the identification task from the prior experiment (with 12 stimuli) for training purposes. Poor performance on this training was used for exclusion purposes. Three people with low accuracy ($6\%$, $6\%$, and $25\%$ accuracy compared to a mean of $70\%$) were excluded.

@@

We had two hypotheses, stemming from our belief that VSUPs promote better \emph{integration} between uncertainty and value information, and encourage \emph{caution} by highlighting the ambiguity introduced by uncertain data. In particular:
\begin{enumerate}
	\item Participants would choose targets with \textbf{lower uncertainty} in the prediction task when using a VSUP. 
	\item This would result in a tradeoff where they would also choose targets with \textbf{lower expected value} when using a VSUP.
\end{enumerate}

@@

\subsubsection{Results} 
\taskTwoViolin
Consistent with our first experiment, we found no significant effect of legend shape on either uncertainty ($F(1,61) = 0.01$, $p=0.92$) or value ($F(1,61)=3.1$, $p=0.08$) of guesses. This result suggests that wedge- and square-shaped legends promote similar patterns of decision-making. Overall, we prefer to employ the wedge-shaped legend for VSUPs, and the square-shaped legend for traditional maps, as it makes the conceptual differences between the two more apparent.

The results fail to support our first hypothesis. We performed a repeated measures ANOVA on our results to measure the effect of VSUP versus standard quantizations, and square- versus wedge-shaped legends, on average uncertainty in bets. We found no significant effect of quantization scheme on average guess uncertainty ($F(1,61)=0.05$, $p=0.83$). This result indicates that there does not appear to be a uniform pattern of risk aversion between the two scale types.

The results support our second hypothesis. An ANOVA for average danger value of guesses found a significant effect of quantization scheme ($F(1,61)=17$, $p<0.01$). A post-hoc t-test confirmed ($t=2.3$, $p=0.02$) that guesses made with VSUPs had significantly higher average danger ($M=0.32$, $SD=0.17$) (and so lower expected value) than traditional maps ($M=0.29$, $SD=0.18$). Since there was no significant difference in average guess uncertainty, and the distribution of danger and uncertainty was identical across conditions, this result suggests that VSUPs led participants to moderate their guesses: avoiding highly uncertain regions, but willing to ``gamble'' more on regions of middling certainty. \figref{fig:taskTwoViolin} shows the comparative distribution of uncertainty in participant guesses.

@@

The results from our initial evaluation of VSUPs are promising, showing that a general audience can make use of uncertainty information in a reasonable way. The way that uncertainty information is presented can have a measurable impact on decision-making. The VSUP strategy of only emphasizing differences when uncertainty is sufficiently low is somewhat analogous to the inferential statistics such as effects tests. VSUPs might therefore be used to promote more caution in judgments, and lead analysts away from spurious signals in data.

@@

Uncertainty, data quality, or confidence values are often considered separately from the data itself, relegated to tooltips or visually distant supplemental charts. We contend, in contrast, that uncertainty information ought to be directly integrated within a shared chart. This integration introduces additional complexity in the design and presentation of data. Value-Suppressing Uncertainty Palettes represent one strategy for dealing with this complexity, by assigning mark properties in a way that supports the disambiguation of values in data where uncertainty is low, but suppresses these judgments when uncertainty is high. This decision of how to allocate visual variables promotes patterns of decision-making that make responsible use of uncertainty information, discouraging comparison of values in unreliable regions of the data, and promoting comparison in regions of high certainty.
